---
title: "P8104 Homework 3"
author: "Jenna Mohammed"
date: '2023-10-11'
output: html_document
---

```{r setup}

library(ggridges)
library(tidyverse)
library(patchwork)
library(dplyr)
library(ggplot2)

library(p8105.datasets)
data("instacart")

library(p8105.datasets)
data("brfss_smart2010")
```

# Problem 1 

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

There are 134 aisles, with fresh vegetables and fresh fruit having the most ordered items.

```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```


# Question 2

Loading in and cleaning the data 

```{r}
view(brfss_smart2010)

brfss_df <- brfss_smart2010 |>
  janitor::clean_names() |>
  filter(
    topic == "Overall Health", 
    response %in% c("Excellent", "Very good", "Good", "Fair", "Poor")
    ) |>
  mutate(
    response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))
  )

view(brfss_df)
```

Determining which states were observed at 7 or more locations

```{r}
states_observed_df <- brfss_df |> 
  filter(year %in% c(2002, 2010)) |>
  select(year, locationabbr, locationdesc) |>
  distinct() |>
    group_by(year, locationabbr) |>
  summarize(n=n()) |>
  filter(n >= 7) |>
  arrange(year, desc(n))
  
view(states_observed_df)
  
# below is the first attempted code for this portion
  #group_by(year, locationabbr) |>
  #summarise(location_count = n_distinct(locationdesc), .groups = "drop") |>

#filter(locationdesc >= 7)|>
  #arrange(desc(locationdesc))

# need help -> says locationdesc not found
```

Construct a dataset

```{r}
# creating a datatset filtering "Excellent" values

Excellent_dataset <- brfss_df |>
  janitor::clean_names() |>
  filter(response == "Excellent")
  
#view(Excellent_dataset)

# averaging the values 

average_df <- Excellent_dataset %>%
  group_by(year, locationabbr) %>%
  summarize(average_df = mean(data_value, na.rm = TRUE))

#view(average_df)

# make a spaghetti plot of the average value over time depending on state

ggplot(average_df, aes(x = year, y = average_df, group = locationabbr, color = locationabbr)) + 
  geom_point() + 
  geom_line() +
  labs( title = "Plot of Average Value Over Time by State",
    x = "Year",
    y = "Average Value")


```

Making a two panel plot in NY data

```{r}
ny_df <- brfss_df %>%
  filter(locationabbr == "NY" & year %in% c(2006, 2010) & response %in% c("Poor", "Fair", "Good", "Very good", "Excellent"))

ggplot(ny_df, aes(y = data_value, x = response ), alpha = 0.7) +
  geom_col(aes(fill = response)) +
    facet_wrap(~year, ncol = 1) +
    labs(title = "Distribution of Data Value by Response Quality in NY State",
       x = "Data Value",
       y = "Density",
       fill = "Response Quality")

# one of the response values are missing in the dual graph
```

# Problem 3 

Load in and tidy data for participants

```{r}
covar_df = read_csv("./nhanes_covar.csv", skip = 4) %>%
  janitor::clean_names() |>
  mutate(sex,sex=ifelse(sex==1,"Male","Female"))|>
  mutate(sex,sex=as_factor(sex))|>
  mutate(education = recode(education, "1" = "Less than high school", "2" = "High school equivalent", "3" = "More than high school")) |>
  filter(age>=21)|>
  drop_na()|>

view(covar_df)
```

load in and tidy data for accelerator

```{r}
accel_df = read_csv("./nhanes_accel.csv") |>
  janitor::clean_names() |>
  drop_na()

view(accel_df)
```

Merge participant and accelerator data

```{r}
combined_data <- left_join(covar_df,accel_df,by= "seqn")

view(combined_data)
```

Producing a reader-friendly table

```{r}
combined_data |>
group_by(sex,education)|>
  summarize(n_people=n())|>
  
  pivot_wider(
    names_from = education,
    values_from = n_people
  )

# creating a plot for the above data 

ggplot(combined_data, aes( x = education, y = sex)
       )

ggplot(combined_data, aes(x = education, y = age, fill = sex)) +
  geom_boxplot() +
  labs(title = "Distribution of Age based on Gender",
       x = "Education Level",
       y = "Age") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

Creating the plot for total activities

```{r}
combined_data |>
   mutate(total_act = rowSums(across(min1:min1440)))|>
  ggplot(aes(x=age,y=total_act, color = sex)) +
  geom_point(size = 1.5)  +
  geom_smooth(se=FALSE) +
  labs(
    y="Total MIMS Activity",
    x = "Age",
    title = " Total Activities vs Age ") +
  viridis::scale_color_viridis(name = "Sex", discrete = TRUE) +
    ggthemes::theme_excel() +
  facet_grid(.~education)

```



To ask TA 
- for the two panel plot in question 2, my data is missing  a response value
